A Gentle Introduction to Compressed Sensing

### 2017-01-14 compressed_sensing/cs.tex
cs: intro is significanlty improved

\section{Motivational introduction}
In the last decade the humanity has come to the point when it produces much more information than it can store and analyse.
There is a huge exposion in the variety of sensors and their dimensionality in many areas: video, audio, scientific, medical research.
Because of that we face the problem of \textbf{data deluge} and naturally come to the necessity of compressing data.

\begin{figure}[H]
    \begin{center}
        \includegraphics[width=.75\textwidth]{figures/CS_sample_compress_paradigm.pdf}
    \end{center}
    \caption{
        \label{fig:CSSampleCompressParadigm}
        ``Sample-Then-Compress Paradigm''.
        Sample could be a digital camera, a microphone,~\ldots\, Compression strategy is signal-dependent.
    }
\end{figure}
Let us consider the so-called ``Sample-Then-Compress Paradigm'' (Fig~\ref{fig:CSSampleCompressParadigm}).
We want to transmit a signal (it could be continuous, it does not matter at this point), but we are unable to do it.
Instead we sample it $n$ times (we will say that the dimension of the sample is $n$).
The transmission channel we reduce the dimension of the sampled signal to $S \ll n$.
Afterwards, the received compressed signal is decompressed to previous dimension $n$.
In this scenario $S$ depends on the nature of the signal.
First, with $S \ll n$ the process seems to be somewhat wasteful: why do we actually need to do $n$ samples if we anyway compress in to the dimension of $S$?!
Secondly, every measurement of total $n$ in a sample could be exorbitant.
Thus, we want to reduce $n$ as much as it is possible.

This example is not accurate as we will see, as follows.
However, it demonstrates why people were concerned about Compressed Sensing in the first place.

\section{Introduction to the Compressed Sensing idea}
The idea of Compressed Sensing is that the signal at Fig.~\ref{fig:CSSampleCompressParadigm} that we would like to transmit/acquire is actually sparse (or ``approximately sparse'') signal in a certain ``basis''.
Suppose that this signal is an $m$-dimensional vector, where $m$ is very large.
Then we want to find a linear basis in $\mathbb R^m$ for this vector to be $S$-sparse (i.e. to have only $S$ nonzero components).
The challenge of the problem is that we do not observe the whole signal.\footnote{
    Suppose that we do.
    Then we can use it as an orth in a basis and the signal is $1$-sparse.
    However, we cannot observe the whole signal because $m$ is very large.
    Take a look at Fig.\ref{fig:CSSampleCompressParadigm}.
    If the signal is an $m$-dimensional vector, then to find the basis we would need to take $n := m$ samples.
    But as we have already discussed we want to reduce $n$ as much as it is possible.
    Because of that we want to find a basis of interest without observing the whole signal.
}

Let us suppose that we now deal with a $S$-sparse signal (we still do not observe it, but know that it is $S$-sparse).
Our aim is to reconstruct the signal (think about decompressing at Fig.\ref{fig:CSSampleCompressParadigm}).
We can make measurements that are a certain transformation of the signal (for example, linear).
We want to make as few measurements as possible (to reduce $n$ at Fig.~\ref{fig:CSSampleCompressParadigm}).
Without the assumption of us possessing the basis in which the original signal is sparse the problem would have a trivial solution: we would generally need to make as many measurements as is the dimension of a signal.\footnote{Think it through!}

@@

In these notes we will not consider the problem of finding the basis in which the signal can be decomposed in a sparse manner.
We will only limit ourselves to a very small example given in the section \ref{sec:motivation} that demonstrates that this assumption is feasible.
Our main concern will be the second problem: to find and quantify the minimal possible number of measurements to reconstruct the original signal.

>>>>

\section{Introduction}
In the last decade the humanity has come to the point when it produces much more information than it can store and analyse.
There is a huge exposion in the variety of sensors and their dimensionality in many areas: video, audio, scientific, medical research.
Because of that we face the problem of \textbf{data deluge} and naturally come to the necessity of compressing data.

The problem that we will consider can be calculated as follows.
We want to transmit a signal (usually continuous), but we are unable to do it, because the transmission channel is quite narrow.
What strategies can we come up with?

Let us consider the so-called ``Sample-Then-Compress Paradigm'' (Fig~\ref{fig:CSSampleCompressParadigm}).
We sample the signal $n$ times (vector in $\mathbb R^n$).
The transmission channel is narrow, thus, we lossy compress the sample to the dimension $d \ll n$ (vector in $\mathbb R^d$).
Afterwards, the received compressed signal is decompressed to previous dimension $n$.

This strategy has a number of flaws.
First, how big should $n$ be for us to recover the signal ``well enough'' and is it possible with the limitation of $d$?
Secondly, with $d \ll n$ the process seems to be somewhat wasteful: why do we actually need many $n$ samples if we anyway compress in to the dimension $d$?!
Thirdly, at which time points should we measure the signal?
Fourthly, every measurement in the sample could be exorbitant.
Thus, we want to reduce $n$ as much as it is possible.

\begin{figure}[H]
    \begin{center}
        \includegraphics[width=.60\textwidth]{figures/CS_sample_compress_paradigm.pdf}
    \end{center}
    \caption{
        \label{fig:CSSampleCompressParadigm}
        ``Sample-Then-Compress Paradigm''.
        Sample could be a digital camera, a microphone,~\ldots\, Compression strategy is signal-dependent.
    }
\end{figure}

\begin{center}
    \bf \color{red} Maybe we can do better?
\end{center}

\section{Compressed Sensing idea}
The idea of Compressed Sensing is that the signal at Fig.~\ref{fig:CSSampleCompressParadigm} that we would like to transmit/acquire is often sparse (or ``approximately sparse'') signal in a certain ``basis''.
We will consider a very small example in the section \ref{sec:motivation} that demonstrates that this assumption is feasible.
Suppose that this signal is an $m$-dimensional vector, where $m$ is very large.
Then we want to find a linear basis in $\mathbb R^m$ for this vector to be $S$-sparse (i.e. to have only $S$ nonzero components).
The challenge of the problem is that we do not observe the whole signal.\footnote{
    Suppose that we do.
    Then we can use it as an orth in a basis and the signal is $1$-sparse.
    However, we cannot observe the whole signal because $m$ is very large.
    Take a look at Fig.\ref{fig:CSSampleCompressParadigm}.
    If the signal is an $m$-dimensional vector, then to find the basis we would need to take $n := m$ samples.
    But as we have already discussed we want to reduce $n$ as much as it is possible.
    Because of that we want to find a basis of interest without observing the whole signal.
}

Let us suppose that we now deal with a $S$-sparse signal (we still do not observe it, but know that it is $S$-sparse).
Our aim is to transmit the signal.
We can make measurements that are a certain transformation of the signal (for example, linear).
We want to make as few measurements as possible (to reduce $n$ at Fig.~\ref{fig:CSSampleCompressParadigm}).
Without the assumption of us possessing the basis in which the original signal is sparse the problem would have a trivial solution:
we would generally need to make as many measurements as is the dimension of a signal.\footnote{Think it through!}
But with the sparsity there is a chance that we need less measurements than the width of the transmission channel $d$
to reconstruct the original signal exactly or approximately.

@@

In these notes we will not consider the problem of finding the basis in which the signal can be decomposed in a sparse manner.
We will only limit ourselves to a very small example given in the section \ref{sec:motivation}.
Our main concern will be the second problem.
